---
output:
  pdf_document: default
  html_document: default
---
# Data Source
The data explored was taken from Kaggle: [Video Game Sales with Ratings](https://www.kaggle.com/rush4ratio/video-game-sales-with-ratings).

# Reading the Data
```{r,results='hide',warning=FALSE,message=FALSE}
#Importing libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(data.table)
library(ggplot2)
library(knitr)
library(stringr)

#Loading the dataset from local storage.
vgsale <- read_csv("C:/Users/achu3/Documents/datasets/Video_Games_Sales_as_at_22_Dec_2016.csv/Video_Games_Sales_as_at_22_Dec_2016.csv")
```
# Data Cleaning & Manipulation
```{r,message=FALSE,warning=FALSE}
str(vgsale)
summary(vgsale)
```
```{r,message=FALSE,warning=FALSE}
#Missing Value Inspection & Filter
vgsale <- vgsale%>%filter(!is.na(Name)) #only clearing character NAs

#Transforming user/critic scores, release years, and similar to numeric format for use.
vgsale1 <- transform(vgsale,User_Score=as.numeric(User_Score),Year_of_Release=as.numeric(Year_of_Release)) #char->num
```

```{r,message=FALSE,warning=FALSE}
vgsale1%>%ggplot(aes(x=Year_of_Release))+
  geom_histogram(stat="count",fill="#0072B2")+ #I like #0072b2 as a colorblind-friendly option.
  labs(title="release year range in dataset")
```
This dataset includes data from Metacritic, such as critic and user scores, 
which was launched in 1999.
Since we will be using its metrics to explore publishers' performance, 
we will filter out games from 1999 to the most recent year covered 
in the dataset (2016).
```{r}
vgsale2 <- vgsale1%>%filter(Year_of_Release>=1999) #I prefer creating new datasets for bigger manipulations in case I need to quickly return to a previous version of the data.
```

```{r,results='hide'}
#Publisher subsidiaries are merged & renamed with parent companies.
levels(factor(vgsale2$Publisher)) #ls unique names;brief googling if unsure about similar names
#Activision*;Codemasters & Codemasters Online appear to be different;EA Games, Electronic Arts*;Enix Corp, Square* (including squaresoft);Idea Factory*;Marvelous*;Rebellion*;Sony*;Ubisoft*;Zoo*, Zushi Games
#Unknown = Unknown Worlds? check game of obsv
vgsale2$Publisher <- sub("^Activision.*","Activision",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Electronic Arts.*","Electronic Arts",vgsale2$Publisher)
vgsale2$Publisher <- sub("EA Games","Electronic Arts",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Square.*","Square Enix",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Enix.*","Square Enix",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Idea Factory.*","Idea Factory",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Marvelous.*","Marvelous Inc",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Rebellion.*","Rebellion",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Sony.*","Sony",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Ubisoft.*","Ubisoft",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Zoo.*","Zoo Games",vgsale2$Publisher)
vgsale2$Publisher <- sub("^Zushi Games.*","Zoo Games",vgsale2$Publisher)

#Sort multiples (ex. Activision - CoD:BOII) by platform.
vgsale2$Name_Platform <- paste(vgsale2$Name,vgsale2$Platform,sep=", ")
```

Exploring - publisher focus.
```{r,message=FALSE,warning=FALSE}
#What are the top ten publishing companies? by global sale
pubsales <- vgsale2%>%group_by(Publisher)%>%
  summarise(pubttlsale=sum(Global_Sales))%>%
  arrange(desc(pubttlsale))%>%
  top_n(10)%>%
  mutate(wrapname=str_wrap(Publisher,width=15))
pubsales%>%ggplot(aes(x=reorder(wrapname,pubttlsale),y=pubttlsale))+
  geom_bar(stat="identity",fill="#0072B2")+
  coord_flip()+
  labs(x="Publisher",y="Total Sales (millions of copies)",
       title="Top 10 Video Game Publishers by Global Sales",subtitle="1999-2016")
```
```{r,message=FALSE,warning=FALSE}
#What are the top ten games by publishing companies? by global sale
topgamesale <- vgsale2%>%top_n(10,Global_Sales)%>%mutate(wrapname=str_wrap(Name,width=15))
topgamesale%>%ggplot(aes(x=reorder(wrapname,Global_Sales),y=Global_Sales))+
  geom_bar(stat="identity",fill="#0072B2",position="dodge")+
  coord_flip()+
  labs(x="Game",y="Total Sales (millions of copies)",
       title="Top 10 Games by Global Sales",subtitle="1999-2016")
#Nintendo dominates in sales.
#Why is that?
```

```{r,message=FALSE,warning=FALSE}
#What genres do publishers create best?

#top 5 games(sales)/top 5 publisher (w/Genre) (top 5 to narrow focus)
#This gives us a brief look as to what type of games do well per publisher.
#ex. Activation is pure shooter, while Nintendo has the most variety.
pubsales_5 <- vgsale2%>%group_by(Publisher)%>%
  summarise(pubttlsale=sum(Global_Sales))%>%
  arrange(desc(pubttlsale))%>%
  top_n(5) #top five publishers from prev top pub subset
topgamesale2 <- vgsale2%>%group_by(Publisher)%>%filter(Publisher%in%pubsales_5$Publisher)%>%
  select(Name_Platform,Publisher,Genre,Global_Sales)%>%
  arrange(desc(Global_Sales))%>%
  top_n(5) #top five games/publisher by sale
topgamesale2 <- topgamesale2%>%mutate(Publisher=reorder(Publisher,-Global_Sales),#order pub by sales
                                      Name_Platform=reorder_within(
                                        Name_Platform,Global_Sales,Publisher,fun=sum)) #order name by sales per pub
topgamesale2%>%ggplot(aes(x=Global_Sales,y=Name_Platform,fill=Genre))+
  geom_col()+ #functions the same as flipped geombar(identity)
  #geom_text(aes(label=Global_Sales),hjust=1)+
  facet_wrap(~Publisher,ncol=1,scales="free_y")+
  scale_y_reordered()+
  labs(x="Total Sales (millions of copies)",y="Game",
       title="Top Five Games per Publisher by Global Sales",subtitle="1999-2016")
```

```{r,message=FALSE,warning=FALSE}
#top 3 genres(sales)/top 5 publisher
#A more pure/diverse look than before.
topgenrepub <- vgsale2%>%filter(Publisher%in%pubsales_5$Publisher)%>%
  select(Publisher,Genre,Global_Sales)%>%
  group_by(Genre,Publisher)%>%
  summarise(genrettlsale=sum(Global_Sales))%>%
  arrange(desc(genrettlsale))%>%
  group_by(Publisher)%>%
  top_n(3)%>%
  mutate(Publisher=reorder(Publisher,-genrettlsale),
         Genre=reorder_within(Genre,genrettlsale,Publisher,fun=sum))
topgenrepub%>%ggplot(aes(x=genrettlsale,y=Genre))+
  geom_col(fill="#0072B2")+
  facet_wrap(~Publisher,ncol=1,scales="free_y")+
  scale_y_reordered()+
  labs(x="Total Sales (millions of copies",y="Genre",
       title="Top 3 Genres per Publisher by Global Sales",subtitle="1999-2016")
#most frequently - sports, shooter, action;
#Nintendo seems to specialize with platforming games; has unique range compared to others
#confirms Activision & shooters; EA dominates sports (in terms of sales)
```

```{r,message=FALSE,warning=FALSE}
####look into the percentage sales per genre per publisher

#Do these sales match with score?
#average user score of games/publisher
topuserscorepub <- vgsale2%>%filter(Publisher%in%pubsales$Publisher)%>%filter(User_Count>=40)%>%
  select(Publisher,User_Score)%>%
  group_by(Publisher)%>%
  summarise(avguserscore=mean(User_Score))%>%
  mutate(wrappub=str_wrap(Publisher,width=15)) 
topuserscorepub%>%ggplot(aes(x=reorder(wrappub,-avguserscore),y=avguserscore))+
  geom_bar(stat="identity",fill="#0072B2")+
  labs(x="Publisher",y="Average User Score",title="Average User Score of Games by Publisher",subtitle="1999-2016")
```
```{r,message=FALSE,warning=FALSE}
#average critic score of games/publisher
topcriticscorepub <- vgsale2%>%filter(Publisher%in%pubsales$Publisher)%>%filter(Critic_Count>=40)%>%
  group_by(Publisher)%>%
  summarise(avgcriticscore=mean(Critic_Score))%>%
  mutate(wrappub=str_wrap(Publisher,width=15))
topcriticscorepub%>%ggplot(aes(x=reorder(wrappub,-avgcriticscore),y=avgcriticscore))+
  geom_bar(stat="identity",fill="#0072B2")+
  labs(x="Publisher",y="Average Critic Score",title="Average Critic Score of Games by Publisher",subtitle="1999-2016")
#Not much to see in terms of publisher score - scores are fairly close together.
```

```{r,message=FALSE,warning=FALSE}
##Platform
#most games on a platform
levels(factor(vgsale2$Platform))
vgsale2%>%ggplot(aes(x=Platform))+
  geom_histogram(stat="count",fill="#0072B2")
```

```{r,message=FALSE,warning=FALSE}
##Genre
vgsale2%>%ggplot(aes(x=Genre))+
  geom_histogram(stat="count",fill="#0072B2") #game/genre
```

```{r,message=FALSE,warning=FALSE}
genresale <- vgsale2%>%group_by(Genre)%>%
  summarise(genrettlsale=sum(Global_Sales))%>%
  arrange(desc(genrettlsale))
genresale%>%ggplot(aes(x=reorder(Genre,genrettlsale),y=genrettlsale))+
  geom_bar(stat="identity",fill="#0072B2")+
  coord_flip()+
  labs(x="Genre",y="Total Sales (millions of copies)",title="Total Video Game Sales by Genre",subtitle="1999-2016")
#but is action, sports rated highly?
#who buys these? what region? region&score genre
```

```{r,message=FALSE,warning=FALSE}
genreuser <- vgsale2%>%filter(User_Count>=40)%>%
  group_by(Genre)%>%
  summarise(avgscore_user=mean(User_Score))%>%
  arrange(desc(avgscore_user))
genreuser%>%ggplot(aes(x=reorder(Genre,avgscore_user),y=avgscore_user))+
  geom_bar(stat="identity",fill="#0072B2")+
  coord_flip()+
  labs(x="Genre",y="Average User Score",title="Averge User Score by Video Game Genre",subtitle="Metacritic, 1999-2016")
```

```{r,message=FALSE,warning=FALSE}
criticuser <- vgsale2%>%filter(Critic_Count>=40)%>%
  group_by(Genre)%>%
  summarise(avgscore_critic=mean(Critic_Score))%>%
  arrange(desc(avgscore_critic))
criticuser%>%ggplot(aes(x=reorder(Genre,avgscore_critic),y=avgscore_critic))+
  geom_bar(stat="identity",fill="#0072B2")+
  coord_flip()+
  labs(x="Genre",y="Average Critic Score",title="Average Critic Score by Video Game Genre",subtitle="Metacritic, 1999-2016")
#seemingly not too much difference w/in groups, although top differs 
```

```{r,message=FALSE,warning=FALSE}
#Highest Rated Games (User Score)
hiscore_user <- vgsale2%>%filter(User_Count>=40)%>%
  select(Name,User_Score)%>%
  arrange(desc(User_Score))%>%
  #top_n(5)
  slice(1:5)
hiscore_user$wrapname <- str_wrap(hiscore_user$Name,width=15)
hiscore_user%>%ggplot(aes(reorder(wrapname,-User_Score),y=User_Score))+
  geom_bar(stat="identity",fill="#0072B2")+
  geom_text(aes(label=User_Score),vjust=-.25)+
  #scale_x_discrete(guide=guide_axis(n.dodge=3))+
  #scale_x_discrete(labels=abbreviate)+
  labs(x="Game",y="User Score",title="Highest Scoring Games by User",subtitle="1999-2016")
```

```{r,message=FALSE,warning=FALSE}
#Highest Rated Games (Critic)
hiscore_critic <- vgsale2%>%filter(Critic_Count>=40)%>%
  select(Name_Platform,Critic_Score)%>%
  arrange(desc(Critic_Score))%>%
  slice(1:5)
hiscore_critic$wrapname <- str_wrap(hiscore_critic$Name_Platform,width=15)
hiscore_critic%>%ggplot(aes(reorder(wrapname,-Critic_Score),y=Critic_Score))+
  geom_bar(stat="identity",fill="#0072B2")+
  geom_text(aes(label=Critic_Score),vjust=-.25)+
  #scale_x_discrete(guide=guide_axis(n.dodge=3))+
  #scale_x_discrete(labels=abbreviate)+
  labs(x="Game",y="Critic Score",title="Highest Scoring Games by Metacritic",subtitle="1999-2016")
#GTA on multiple platforms...maybe instead of combining names, average the
#scores for multiples?
```

